---
title: "Change of Mind Analyses"
author: "Anonymised for peer review"
format: 
  pdf: 
    execute:
      echo: false
      warning: false
toc: true
number-sections: true
bibliography: references.bib
jupyter: python3
---

# Was the original stance against the claim?

```{r}
load(here("data/clean/all_data_clean.Rdata"))
data <- all_data$data
```

```{r}
# Define a function for plotting a basic grouped bar plot with optional ordering
plot_grouped_bar <- function(by_variable, column, x_label, usage_ordering = NULL, label_angle = 0, show_n = FALSE) {
  colour_scale <- c("purple", "darkblue", "red")
  
  # Check if usage_ordering is provided and build the base plot accordingly
  if (!is.null(usage_ordering)) {
    plot <- ggplot(data = by_variable, aes(x = factor({{column}}, levels = usage_ordering), y = mean_update, fill = consensus)) +
      geom_bar(position = "dodge", stat = "identity", color = "black", alpha = .8) + 
      geom_errorbar(aes(ymin = mean_update - se, ymax = mean_update + se), 
                    position = position_dodge(0.9), width = 0.25) + 
      labs(x = x_label, y = "Mean Update (Post - Prior)") + 
      theme_minimal() + 
      theme(axis.text.x = element_text(angle = label_angle, hjust = 1))
  } else {
    plot <- ggplot(data = by_variable, aes(x = {{column}}, y = mean_update, fill = consensus)) +
      geom_bar(position = "dodge", stat = "identity", color = "black", alpha = .8) + 
      geom_errorbar(aes(ymin = mean_update - se, ymax = mean_update + se), 
                    position = position_dodge(0.9), width = 0.25) + 
      labs(x = x_label, y = "Mean Update (Post - Prior)") + 
      theme_minimal() + 
      theme(axis.text.x = element_text(angle = label_angle, hjust = 1))
  }

  # Add color and fill scales
  plot <- plot + 
    scale_color_manual(values = colour_scale) +
    scale_fill_manual(values = colour_scale)+
    theme(legend.position = "bottom")
  
  return(plot)
}

```

```{r}
by_original_stance_claim <- data %>%
  group_by(consensus, original_stance_claim) %>%
  summarise(
    mean_update = mean(update),
    n_obs = n(),
    se = sd(update) / sqrt(n_obs)
  )


plot_original_stance_claim <- plot_grouped_bar(by_original_stance_claim, original_stance_claim, x_label = "Original Stance on Claim")
plot_original_stance_claim + geom_text(data = by_original_stance_claim, aes(label = n_obs, x = original_stance_claim, y = -5), position = position_dodge(width = 1))

ggsave(here("analyses/07_Plots/original-stance-claim.png"), width = 7, height = 3)

```

# On how many trials did people qualitatively changed their beliefs  (change their minds)

## go from disbelieving -\> believing/believing -\> disbelieving

```{r}
colour_scale <- c("purple", "blue", "red")

data %>%
  filter(!is.na(changed_mind)) %>%
  group_by(consensus, changed_mind, claim_type) %>%
  summarise(n_trials = n()) %>%
  ggplot(aes(x = changed_mind, y = n_trials, fill = consensus, group = consensus))+
  geom_col(position = "dodge", alpha = .8)+
  facet_wrap(~claim_type, nrow = 1)+
  labs(y = "Number of Trials", x = "Changed Mind")+
  scale_fill_manual(values = colour_scale)+
  theme_bw()+
  theme(legend.position = "bottom")
ggsave(here("analyses/07_Plots/change-mind.png"), width = 7, height = 3)


```

People did not change their mind in the majority of trials. Unsurprisingly, people changed their minds most in the consensus trials, and there didn't seem to be much of a difference between independent versus dependent.

```{r}
d_change_mind <- data %>%
  filter(!is.na(changed_mind)) %>%
  group_by(consensus, changed_mind) %>%
  summarise(n_trials = n()) %>%
  ungroup() %>%
  mutate(percent = round((n_trials/sum(n_trials))*100,2))

d_change_mind_claim <- data %>%
  filter(!is.na(changed_mind)) %>%
  group_by(claim_type, changed_mind) %>%
  summarise(n_trials = n()) %>%
  ungroup() %>%
  mutate(percent = round((n_trials/sum(n_trials))*100,2))

d_change_mind
sum(d_change_mind$percent[d_change_mind$changed_mind == TRUE])
sum(d_change_mind$percent[d_change_mind$changed_mind == TRUE & d_change_mind$consensus != "contested"])
```

